---
title: "401 HW 6"
output: html_document
---
1.) 

```{r}
setwd("~/Desktop")
data <- read.table(file = "senic.txt", header = TRUE)



model <- lm(data[[2]] ~ data[[4]] + data[[3]] + log(data[[7]]) + data[[9]])
plot(data$LengthOfStay,model$residuals); abline(h=0)

 qqnorm(model$residuals); qqline(model$residuals)
```


The diagnostic plots look slightly suspicious for violations of Gaussian properties, as there seems to be a positive correlation on the residual plot and some deviation on the qq plot.


2a.)
``` {r}
model2 <- lm(data[[2]] ~ data[[4]] + data[[3]] + log(data[[7]]) + data[[9]])
confint(model2)


```

2b.)
``` {r}

a <- data[[9]]
for ( itr in 1:113 ){
  if(data[itr, 9] >1){
    a[itr] = 0
  }
}

b <- data[[9]]
for ( itr in 1:113 ){
  if(data[itr, 9] <2){
    b[itr] = 0
  }
}

model3 <- lm(data[[2]] ~ data[[4]] + data[[3]] + log(data[[7]]) + data[[8]] + a*b)
summary(model3)
confint(model3)

```

2c.)
``` {r}

model4 <- lm(data[[2]] ~ data[[3]] + log(data[[7]]) + data[[9]] + data[[4]]*data[[8]])
summary(model4)
confint(model4)
```



3a.)

``` {r}
model5 <- lm(data[[2]] ~ data[[4]] + data[[10]] + log(data[[7]]) + data[[9]])

plot(data$LengthOfStay,model5$residuals); abline(h=0)
qqnorm(model5$residuals); qqline(model5$residuals)

```


``` {r}
x = data[[4]]
model6 <- lm(data[[2]] ~ data[[4]] + poly(x,2) + data[[10]] + log(data[[7]]) + data[[9]])

summary(model6)

plot(data$LengthOfStay,model6$residuals); abline(h=0)
qqnorm(model6$residuals); qqline(model6$residuals)


```

The models appear to have about the same fits when looking at the diagnostic checks. Perhaps the model with the log term has a slightly better fit, but overall they appear to have roughly the same level of fit.

