---
title: "Untitled"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
Sys.getenv("RSTUDIO_PANDOC")
```


2.)


```{r}
setwd("~/Desktop")
data <- read.csv(file = "bea-2006 (1).csv", header = TRUE)

x <- log(data[[3]])
y <- data[[2]]

model <- lm(formula = y ~ x, data=data)
summary(model)
model_summ <- summary(model)
mse <- mean(model_summ$residuals^2)
mse
condvar <- mean(residuals)^2
condvar
```



2c.) Pittsburgh has a population of 2361000, and a per capita GMP of 38350. For model prediction see below:

```{r}

y <- log(2361000)*4449.8 - 23306.2
y

residual <- 38350 - y

residual

```

Predicted value is 41992.82, and the residual is -3642.816

```{r}
setwd("~/Desktop")
data <- read.csv(file = "bea-2006 (1).csv", header = TRUE)

x <- log(data[[3]])
y <- data[[2]]

model <- lm(formula = y ~ x, data=data)

residuals <- resid(model)

plot(y, residuals)
abline(0,0)

```


If the assumptions of a simple linear model were to hold, there should be no pattern to the residual plot. Here there appears to be a linear pattern to the residual plot, suggesting that a linear model does not hold.


2e.)
Estimated Slope
```{r}
slope <- 4449.8
error <- 390.9
l <- slope - 1.645*error
h <- slope + 1.645*error
l
h



```



Estimated slope interpretation: For 2 values whose log populations happen to differ by 1, the expected difference in GMP is 4449.8.



3.) 

```{r}
setwd("~/Desktop")
data <- read.table(file = "bp.txt", header = TRUE)

data


plot(x = data[[1]], y = data[[2]])
model <- lm(formula = data[[2]] ~ data[[1]], data = data)
summary(model)

abline(model, color="blue", lwd=3)

```

Estimate for person who is 200 pounds:

```{r}
estimate <- 200*.41942 + 69.10437
estimate

```

Confidence interval:
```{r}
l <- 152.9884 - 1.96*sqrt(var(data[[2]]))
l
h <- 152.9884 + 1.96*sqrt(var(data[[2]]))
h
```


The 95% C.I is 126.6859, 179.2909

Interpretation: The best linear model prediction for the blood pressure of a person that weighs 200 pounds is 152.9984. We are 95% sure that a person that weighs 200 pounds would have a true mean blood pressure in the range 126.6859, 179.2909.


3b.) New person 145 pounds. The estimated blood pressure and C.I:

```{r}
guess <- 145*.41942 + 69.10437
guess
l <- guess - 1.96*sqrt(var(data[[2]]))
l
h <- guess + 1.96*sqrt(var(data[[2]]))
h
```


Estimate: 129.9203
C.I.: 103.6178, 156.2227






