---
title: "401HW4"
output: html_document
---


```{r}
setwd("~/Desktop")
data <- read.table(file = "bp.txt", header = TRUE)

t <-qt(.975, df=24)
model <- lm(formula = data[[2]] ~ data[[1]], data = data)
summary(model)

residuals <- resid(model)
v <- sum(residuals^2)/24
x <- 200*.41942 + 69.10437
x

sePred <- v*(1 + ((x - mean(data[[1]]))^2)/var(data[[1]])/26)

l = x-t*sePred
h = x+t*sePred
l
h
```

The prediction interval predicts in what range a future individual observation will fall, while a confidence interval shows the likely range of values associated with some statistical parameter of the data. The prediciton interval here is wider numerically.

1b.) 
Since 140 is in our 95% prediction interval, the data does not allow us to reject this hypothesis. The p value is,

```{r}
z <- (152.9884-140)/sePred
z
1-pnorm(z)
```
The p value is .435



2c.)
```{r}
A <- matrix(c(1,0,1,1,0,1), nrow=2, ncol=3, byrow = TRUE)
w <- matrix(c(1,0,.5,0,1,0,.5,0,1), nrow=, ncol=3, byrow = TRUE)
A%*%w%*%t(A)

```


3b.)

```{r}
x <- matrix( c(1, 9, 3, 6, 1, 25, 8, -16, 1, -3, -27, 8, 1, -10, 4, 3, 1, 1, -1, 1), nrow=5, ncol=4, byrow=TRUE)
m <- t(x)%*%x 
m
solve(m)


```

3c.)
```{r}
x <- matrix( c(1, 9, 3, 6, 1, 25, 8, -16, 1, -3, -27, 8, 1, -10, 4, 3, 1, 1, -1, 1), nrow=5, ncol=4, byrow=TRUE)
y <- matrix( c(12, -30, 13, 8, 4), nrow=5, ncol=1, byrow=TRUE)

b <- solve(m)%*%t(x)%*%y
b
model <- lm(formula = y ~ x)
summary(model)
```
The estimates are the same.


3d.)
```{r}
h <- x%*%solve(m)%*%t(x)
h


```



